---
title: "Election Forecasting"
author: "Roey Stern"
date: '2022-03-29'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(maps)
library(ggmap)
```

```{r}
poll <- read.csv('Polling.csv')
glimpse(poll)
```

```{r}
StatesMap = map_data("state")
electionMap =  merge(StatesMap, poll, by.x = "region",
by.y = "State")
electionMap =electionMap[order(electionMap$order),]
```

```{r}
ggplot(electionMap, aes(x = long, y = lat,fill = DiffCount,group=group)) + geom_polygon() + scale_fill_gradient(low = "blue", high = "red")
```

* Florida looks like it has a very clear Republican presence. 

```{r}
ggplot(electionMap, aes(x = long, y = lat,fill = Rasmussen,group=group)) + geom_polygon() + scale_fill_gradient(low = "blue", high = "red")
```

```{r}
ggplot(electionMap, aes(x = long, y = lat,fill = SurveyUSA,group=group)) + geom_polygon() + scale_fill_gradient(low = "blue", high = "red")
```

* There are a lot more NAs in the two other graphs.
* Some states get a different color in different plots. 
* All the plots show Florida going Republican when it ended up democratic. 

```{r}
ggplot(electionMap, aes(x = long, y = lat,fill = Republican,group=group)) + geom_polygon() + scale_fill_gradient(low = "blue", high = "red")
```

* If I had to create a predictive model I would use the `DiffCount` variable since it has more states predicted and looks more like the actual results. 
