---
title: "Linear Regression HW"
author: "Roey Stern"
date: "2/8/2022"
output:
  
  html_document: default
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Libraries
```{r}
library(tidyverse)
```

### The Data
```{r}
StateData <- read_csv("StateData.csv")
summary(StateData)
```
#### a)i) Scatterplot
```{r a1}
ggplot(data=StateData, aes(x=Longitude, y=Latitude)) + geom_point(color= "Blue") 
```
#### a)ii) Mean High School Grad by Region
```{r a2}
mean_by_region <- StateData %>%
  group_by(Region) %>%
  summarize(Mean=mean(HighSchoolGrad)) %>% arrange(desc(Mean))
mean_by_region

```
#### a)iii) Box plot for 'Murder' by Region
```{r a3}
ggplot(StateData, aes(x=Murder, color = Region)) + geom_boxplot() +
  facet_wrap(vars(Region), scales = 'free') +
  theme(legend.position = 'none')
```

#### b) Linear Regression Model

```{r model}
model = lm(data = StateData,LifeExp ~ Population + Income + Illiteracy + Murder + HighSchoolGrad + Frost + Area) 
summary(model)
```
#### b)i) Linear Regression Equation:

##### y = 20.28 + 9.08(Population) - 10.92(Income) + 7.19(Illiteracy) - 9.18(Murder) +11.30(HighSchoolGrad)-18.59(Frost)-28.06(Area)  
#### b)ii) <u>Interpenetration of Income Coefficient:</u>  
The Income coefficient is −10.92585439. Therefore each one unit increase in Income is associated with a -10.92 unit decrease in Life Expectancy.   

#### b)iii) <u>Life Expectancy by Income Plot</u>
```{r plot}
library(ggpubr)
LifeEXP_Income <- StateData %>%
  ggplot(aes(x= Income, y= LifeExp)) + 
  geom_smooth(method="lm") +geom_point() +
  stat_regline_equation() + stat_cor(aes(label=..rr.label..),label.x = 3100 , label.y = 73)
LifeEXP_Income
         
```
  
Based on the scatter plot, 1 unit increase in income is associated with a 0.00074 increase in LifeExp. This is completely different than the coefficient from my model. I beleive the reason for the discrepancy here is that in the model the Income variable was part of a multiple regression with other values and the combination with them affected its coefficient.  
  
#### c)Refining the Model
```{r}
model_2 <- lm(data = StateData, LifeExp ~ Population + Murder + HighSchoolGrad + Frost)
summary(model_2)
```
#### <u>New Equation:</u>

#### y = 20.30 + 8.62(Population) - 9.15(Murder) + 10.66(HighSchoolGrad) - 19.15(Frost) 
  
#### Comparison Between Model 1 and Model 2:
Starting with the R squared values. Model 1 had an R squared of 0.7362 and an adjusted R squared of 0.6922. Model 2 had an R squared of 0.736 and an adjusted R squared of 0.7126. The R squared value was the practically the same between the 2 models. The adjusted R squared has increased in model 2, which indicates that model 2 is stronger and will be better used to predict data.  
The coefficients of the variables that appear in both models have all decreased in their respective manner but 'frost' which grew a bit in model 2. 

### Longitutde and Latitude to States
```{r}
library(sf)
library(spData)
long_lat <- StateData %>%
  select(Longitude, Latitude)
lonlat_to_state <- function(long_lat,
                            states = spData::us_states,
                            name_col = "NAME") {
    pts <- st_as_sf(long_lat, coords = 1:2, crs = 4326)
    
    states <- st_transform(states, crs = 3857)
    pts <- st_transform(pts, crs = 3857)
     state_names <- states[[name_col]]
    ii <- as.integer(st_intersects(pts, states))
    state_names[ii] }

State_names <- lonlat_to_state(long_lat)

StateData <- StateData %>%
  mutate('State' = State_names) 
```
### Predicting
```{r}
Prediction <-predict(model_2, newdata = StateData)

Values <- data.frame(State=StateData$State, Actual_LifeExp =StateData$LifeExp
                           , Predicted_LifeExp = Prediction)
Values

```
#### Analysis - Lowest Life Expectancy
```{r}
lowest_predicted <- filter(Values, Predicted_LifeExp == min(Predicted_LifeExp))
lowest_predicted

lowest_actual <- filter(Values, Actual_LifeExp == min(Actual_LifeExp))
lowest_actual
```
#### Highest Life Expectancy
```{r}
highest_predicted <- filter(Values, Predicted_LifeExp == max(Predicted_LifeExp))
highest_predicted

highest_actual <- filter(Values, Actual_LifeExp == max(Actual_LifeExp))
highest_actual
```


