---
title: "Network - Visualization"
author: "Roey Stern"
date: '2022-03-28'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(tidyverse)
library(igraph)
```

```{r}
users <- read.csv('Users.csv')
glimpse(users)
edges <- read.csv('Edges.csv')
glimpse(edges)
```

```{r}
number_of_users <- users %>% nrow()
avg_friends <- round(nrow(edges)/number_of_users,2)
common_locale <- users %>% filter(School != "") %>% count(Locale)
common_locale
schools <- users %>% count(Gender, School)
schools
```

* The number of users is `r number_of_users`.
* The average number of friends per user is `r avg_friends`. (We have 146 friendships of 59 users.)
* Out of the users that listed a school, the most common locale is B with 19 users.
* No, both School A and School AB have both genders. 

```{r}
g = graph.data.frame(edges, FALSE, users)
plot(g, vertex.size= 5, vertex.label = NA)
```


* There are 3 connected components. 
* There are 7 users with no friends.


```{r}
degree(g)
```


* There are 7 users with 10+ friends. 


```{r}
V(g)$color[V(g)$Gender == "A"] = "red"
V(g)$color[V(g)$Gender == "B"] = "gray"
plot(g, vertex.size = 5, vertex.label = NA)
```


* The coloration shows that there are a lot more B gender and they are friends with more B gender than A gender. 


```{r}
V(g)$color[V(g)$School == "A"] = "red"
V(g)$color[V(g)$School == "AB"] = "gray"
plot(g, vertex.size = 5, vertex.label = NA)
```

* yes it seems from looking at the graph that the AB users (gray) are connected to each other. 


```{r}
V(g)$color[V(g)$Locale == "A"] = "red"
V(g)$color[V(g)$Locale == "B"] = "gray"
plot(g, vertex.size = 5, vertex.label = NA)
```

* This coloration is very clear in this graph and shows that people from the same locale are friends with each other almost exclusively. 
